Lifting CDCL to Relational Domains

### 2016-10-15 domains.tex
complement properties

\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
\end{definition}

For a program with $N$ variables, let $L$ be the total number of 
meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
of all octagonal inequalities in the set $L$.
%
%========================
\input{literals-clauses}
%========================

\subsection{Properties of Domain Elements for Learning}
%
\pscmt{TODO: Review by Peter}
An important property of a clause learning SAT solver 
is that each element of the partial assignments domain 
are decomposable into precisely complementable 
elements.  This property of the domain helps to guide 
the model search away from the conflicting region of the 
search space.

Most numerical abstract domains, such as intervals, octagons, 
polyhedra lacks precise complements, but they can be 
represented as intersections of complementable half-spaces, 
each of which have precise complements.  For example, intervals 
lack complements, but the decomposition of intervals into 
meet-irreducibles have complements.  
\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]
Similarly, the complementation of the octagon in Figure~\ref{octagon} 
can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

Standard abstract interpretation does not require complementation property.  
Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
for abstract interpretation.  We implement a template-based polyhedra domain 
that returns precise complements by decomposing it into half-spaces.

\Omit {
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}. 
}
\Omit {
To this end, we implement our own template-based polyhedra domain and interval 
domain which supports complementation operator.  
%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
  meet-irreducible. It's trivial to do that with APRON. The reason was
a different one: APRON does not support all C operators, e.g. the bitwise
operators.} 
}

\subsection{Static Analysis Equations for Safety}
Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
perform safety analysis by computing fixed point to infer invariants 
over program variables.  However, bounded model checking (BMC) tries to search 
for a counterexample in a bounded execution trace, by symbolically executing a 
program up to a finite bound.  
Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
For a set $N$ of program analysis constraints defined over a set 
of constraint variables $Var = \{X_i| i \in N\}$, representing a program
$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
and the \rmcmt{final valuation} of constraint variables.
\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \]

>>>>

\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
\end{definition}
For example, the meet decomposition of an interval is given as follows.
\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]

For a program with $N$ variables, let $L$ be the total number of 
meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
of all octagonal inequalities in the set $L$.
%
%========================
%\input{literals-clauses}
%========================

\subsection{Properties of Domain Elements for Learning}
%
An important property of a clause learning SAT solver 
is that a partial assignments domain can be decomposed 
into a set of precisely complementable singleton assignments.  
This property of the partial assignments domain helps to 
guide the model search away from the conflicting region 
of the search space and allows case-based analysis.

The precise complementation property of partial assignment lattice 
can be generalized to other lattice structures.  Most numerical 
abstract domains, such as intervals, octagons, polyhedra, can be 
decomposed into half-spaces, that admits precise complements.  
For example, intervals lack complements, but the meet decomposition 
of intervals into meet-irreducibles have precise complements.  
Numerical abstract domains that admit complementable decomposition 
are shown in Figure~\ref{complement}.  Our template-based domain 
implementation precise complements a meet irreducible in an abstraction.
% 
\begin{table}
\begin{center}
{
\begin{tabular}{l l}
\hline
Domain & Precise Complements
Interval & $(x \leq n) \quad \longrightarrow (x > n)$ \\ \hline
Octagon & $(x+y < 1) \quad \longrightarrow (-x-y < 0)$ \\ \hline
Equality & $(x==y) \quad \longrightarrow (x \neq y)$ \\ \hline 
\end{tabular}
\label{fig:complement}
%\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
\end{figure}
%
\Omit{
Similarly, the complementation of the octagon in Figure~\ref{octagon} 
can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]
Standard abstract interpretation do not require complementation property
The application of set-complements on over-approximating 
domains does not give necessary invariants required to reason about assertions
in programs. 
Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
for abstract interpretation.  
}
\Omit {
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}. 
}
\Omit {
To this end, we implement our own template-based polyhedra domain and interval 
domain which supports complementation operator.  
%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
  meet-irreducible. It's trivial to do that with APRON. The reason was
a different one: APRON does not support all C operators, e.g. the bitwise
operators.} 
} 
\Omit {
\subsection{Static Analysis Equations for Safety}
Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
perform safety analysis by computing fixed point to infer invariants 
over program variables.  However, bounded model checking (BMC) tries to search 
for a counterexample in a bounded execution trace, by symbolically executing a 
program up to a finite bound.  
Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
For a set $N$ of program analysis constraints defined over a set 
of constraint variables $Var = \{X_i| i \in N\}$, representing a program
$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
and the \rmcmt{final valuation} of constraint variables.
\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \] 
}
